---
title: "Stats HW11"
author: "Steve Bramhall"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(MASS)    # for studentized residuals
```

####2ai i.	A scatterplot with the following included on the graph: regression line, confidence intervals of the regression line, and prediction intervals of the regression line.
```{r}
# Read data from csv file
year <- c(1992,1994,1996,1998,2000)
prevalence <- c(3.5,5.3,7.8,11.8,18.3)
prevalence = log(prevalence)

data1=data.frame(year,prevalence)

# get min/max values for newx range
summary(data1$prevalence) 
newx <- seq(1992, 2000, by=2)
newy <- seq(0, 20, by=5)

# linear model
linear.regress.model <- lm(prevalence ~ year, data=data1)

# plot linear model
plot(data1$year,data1$prevalence, xlab="Year", ylab="Log-Prevalence", main="Regression")
abline(linear.regress.model, col="lightblue")

# add confidence interval lines
conf.interval <- predict(linear.regress.model,newdata=data.frame(Year=newx),interval="confidence",level = 0.95)
lines(newx, conf.interval[,2], col="blue", lty=2)
lines(newx, conf.interval[,3], col="blue", lty=2)

# add prediction interval lines
pred.interval <- predict(linear.regress.model,newdata=data.frame(Year=newx),interval="prediction",level = 0.95)
lines(newx, pred.interval[,2], col="orange", lty=2)
lines(newx, pred.interval[,3], col="orange", lty=2)
```

####2aii A scatterplot of residuals.
```{r}

resids=resid(linear.regress.model)                       # get the residuals
plot(data1$year,resids,                                  # plot the residuals
     main="Residual Plot: Log-Prevalence vs Year",       # add plot title
     xlab="Year",                                        # add x-axis label
     ylab="Log-Prevalence")                              # add y-axis label

studentized.resids=studres(linear.regress.model)
plot(data1$year,studentized.resids,                      # plot the studentized residuals
     main="Studentized Residual Plot: Log-Prevalence vs Year",# add plot title
     xlab="Year",                                        # add x-axis label
     ylab="Log-Prevalence")                              # add y-axis label

```

####2aiii A histogram of residuals with the normal distribution superimposed.
```{r}
hist(resids,prob=TRUE,
     main="Normal Curve Over Residuals Histogram",       # add plot title
     xlab="Year",                                        # add x-axis label
     ylab="Log-Prevalence")                              # add y-axis label
curve(dnorm(x,mean=mean(resids),sd=sd(resids)),          # add normal curve
      col="darkblue",lwd=2,add=TRUE,yaxt="n")

hist(studentized.resids,prob=TRUE,
     main="Normal Curve Over Studentized Residuals Histogram", # add plot title
     xlab="Year",                                        # add x-axis label
     ylab="Log-Prevalence",                              # add y-axis label
     xlim=c(-4,4))
curve(dnorm(x,mean=mean(studentized.resids),sd=sd(studentized.resids)), # add normal curve
      col="darkgreen",lwd=2,add=TRUE,yaxt="n")
```


####2aiv Once a reasonable model is found (possibly using a transformation), provide a table showing the t-statistics and p-values for the significance of the regression parameters β_(0 ) and β_1.
```{r}
summary(linear.regress.model)
```



