---
title: "Unit12 HW"
author: "Steve Bramhall"
date: "November 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ggplot2)
library(reshape2)
```


```{r}
# Reading Data File
brainDF <- read.csv(".\\Brain.csv",header=TRUE,sep=",",
                       stringsAsFactors = TRUE, encoding = "UTF-8") # read data file
names(brainDF) <- c("Species","Brain","Body","Gestation","Litter")  # rename table columns
```

### Step 1: Review Assumptions
```{r}
Litter=brainDF$Litter  # Use for Legend
ggplot(brainDF,aes(x=brainDF$Gestation,y=brainDF$Brain,col=Litter))+geom_point()+
  labs(title = "Brain Weight vs Gestation",x="Gestation (days)",y="Brain Weight (g)")      # add axes titles
```

### Step 2: Build a Model

```{r}
brainDF$logGestation <- log(brainDF$Gestation)
brainDF$logLitter <- log(brainDF$Litter)
brainDF$logBrain <- log(brainDF$Brain)

fit1 <- lm(brainDF$Brain ~ brainDF$Gestation + brainDF$Litter, data=brainDF)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(fit1)
```

#### The residuals are not random so we need to do a transformation.

```{r}
fit2 <- lm(brainDF$Brain ~ brainDF$logGestation + brainDF$Litter, data=brainDF)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(fit2)
```

#### The residuals are still not random so we need to try a different transformation.

```{r}
fit3 <- lm(brainDF$Brain ~ brainDF$Gestation + brainDF$logLitter, data=brainDF)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(fit3)
```

#### The residuals are still not random so we need to try a different transformation.

```{r}
fit4 <- lm(brainDF$logBrain ~ brainDF$logGestation + brainDF$Litter, data=brainDF)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(fit4)

fit4.resids <- resid(fit4)            # store residuals of model
fit4.student.resids <- rstudent(fit4) # store studentized residuals of model

layout(matrix(c(1,2),1,2)) # optional 2 graphs/page 
hist(fit4.resids,main="Histogram of Residuals",xlab="Residuals")
hist(fit4.student.resids,main="Historgram of Studenized Residuals",xlab="Studenized Residuals")
```

#### The residuals appear random and the residual QQ plot appears linear. 

We will use this model. Log(Brain) = Beta0 + Beta1*Log(Gestation) + Beta2*Litter

### Step 3: Fit the Model
```{r}
summary(fit4) # show results
```

#### Since Litter is not significant, we will remove Litter and re-run the model.

```{r}
fit5 <- lm(brainDF$logBrain ~ brainDF$logGestation, data=brainDF)
summary(fit5)
```


### Step 4: Check Appropriateness of Model
```{r}
ggplot(brainDF,aes(x=log(brainDF$Gestation),y=log(brainDF$Brain),col=Litter))+geom_point()+
  labs(title = "Log(Brain Weight) vs Log(Gestation)",x="Log(Gestation)",y="Log(Brain Weight)")      # add axes titles

layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(fit5)

```

### Step 5: Interpret Coefficients



